Prefix-free graphs and suffix array construction in sublinear space

### 2023-06-24 manuscript/main.tex
added description to figures

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    A recent paradigm shift in bioinformatics from a single reference genome to a pangenome brought with it several graph structures.
    These graph structures must implement operations, such as efficient construction from multiple genomes and read mapping.
    Read mapping is a well-studied problem in sequential data, and, together with data structures such as suffix array and Burrows-Wheeler transform, allows for efficient computation.
    Attempts to achieve comparatively high performance on graphs bring many complications since the common data structures on strings are not easily obtainable for graphs.
    In this work, we introduce prefix-free graphs, a pangenomic data structure allowing us to obtain the well-known data structures from stringology in graph settings, allowing for many efficient operations on pangenomes.
\end{abstract}

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
A recent paradigm shift in bioinformatics from a single reference genome to a pangenome brought with it several graph structures.
These graph structures must implement operations, such as efficient construction from multiple genomes and read mapping.
Read mapping is a well-studied problem in sequential data, and, together with data structures such as suffix array and Burrows-Wheeler transform, allows for efficient computation.
Attempts to achieve comparatively high performance on graphs bring many complications since the common data structures on strings are not easily obtainable for graphs.
In this work, we introduce prefix-free graphs, a novel pangenomic data structure; we show how to construct them and how to use them to obtain well-known data structures from stringology in sublinear space, allowing for many efficient operations on pangenomes.
\end{abstract}

### 2023-06-24 manuscript/pfg.tex
started sa construction
added description to figures

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{images/prefixfree_proof.png}
    \caption{Visual proof that the trigger words induce prefix-free segments.}
    \label{fig:proof}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{images/pfg_gfa.v3.png}
    \caption{Prefix-free graph in GFA format. Link lines omitted for brevity.}
    \label{fig:gfa}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{images/prefixfree_proof.png}
    \caption{
        A sketch of a visual proof that the trigger words induce prefix-free segments.
        Suppose a segment is a prefix of another segment.
        By construction, it ends with a trigger word.
        This trigger word would be in the middle of another segment and would break it into two smaller segments.
    }
    \label{fig:proof}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{images/pfg_gfa.v3.png}
    \caption{
        Prefix-free graph of the running example after normalization and its representation in GFA format.
        Link lines omitted for brevity.
    }
    \label{fig:gfa}
\end{figure}

### 2023-06-24 manuscript/sa_construction.tex
added description to figures

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{images/ids_and_positions.png}
    \caption{
        Segment IDs and segment positions of a segment join. 
        Permuting these according to the inverse permutation of a suffix array results in segment ID array and segment position array shown in Table \ref{tab:suffix}.
    }
    \label{fig:ids_and_positions}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{images/path_join.png}
    \caption{}
    \label{fig:path_join}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{images/ids_and_positions.png}
    \caption{
        Segment join of segments from the running example, corresponding segment IDs, and positions.
        Permuting IDs and positions according to the inverse permutation of the suffix array results in the segment ID array and the segment position array shown in Table \ref{tab:suffix}.
    }
    \label{fig:ids_and_positions}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{images/path_join.png}
    \caption{
        Path join of paths from the running example and the corresponding starts and ranks. The pangenome and segments are shown above to clarify the meaning of the start values.
    }
    \label{fig:path_join}
\end{figure}
